---
title: "Football Analytics"
output: html_notebook
---

# **Installing packages**

-   StatsBomb - For match event data

-   soccermatics - Pitch plots

-   ggsoccer - Pitch plots

```{r}
install.packages("ggsoccer")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("here")

install.packages("devtools")
install.packages("remotes")

devtools::install_github("cran/SDMTools")
devtools::install_github("statsbomb/StatsBombR")
devtools::install_github("jogall/soccermatics")


```

```{r}
library(ggsoccer)
library(tidyverse)
library(here)
library(ggplot2)
library(StatsBombR)
library(soccermatics)
```

# **Data and Pre-processing**

## View all the competitions available in the StatsBomb dataset (Free edition)

```{r}
FreeCompetitions()
```

## Selecting FIFA World Cup 2022

```{r}
comps <- FreeCompetitions() %>%
filter(competition_id==43 & season_name=="2022")
comps
```

## All the match event data available for this tournament

```{r}
matches <- FreeMatches(Competitions = comps)
matches
```

## Choosing the World Cup Finals 2022

```{r}
events <- get.matchFree(matches[10,])
events
```

## Join all the data for this match

```{r}
cleanevents = allclean(events)
cleanevents
```

## Identify player

```{r}
# Converting player name for representation purpose
cleanevents <- cleanevents %>%
  mutate(player.name = ifelse(player.name == 'Lionel Andrés Messi Cuccittini', 'Lionel Messi', player.name))

# Player credentials
player_id <- 5503
player_name <- 'Lionel Messi'
```

# Passes

## Pre-processing - Retrieving location based on period

```{r}
process_period <- function(half) {
  filtered_events <- cleanevents %>%
  filter(type.name == 'Pass' & period==half & player.id==player_id) %>%
  select(location, pass.end_location)

  locationX <- lapply(filtered_events$location, function(x) x[[1]])
  locationY <- lapply(filtered_events$location, function(x) x[[2]])
  endLocationX <- lapply(filtered_events$pass.end_location, function(x) x[[1]])
  endLocationY <- lapply(filtered_events$pass.end_location, function(x) x[[2]])

  filtered_passes <- data.frame(
    LocationX = unlist(locationX),
    LocationY = unlist(locationY),
    EndLocationX = unlist(endLocationX),
    EndLocationY = unlist(endLocationY)
  )

  return (filtered_passes)
}
```

## First Half (Normal time)

```{r}
filtered_passes <- process_period(1)

to_statsbomb <- rescale_coordinates(from = pitch_opta, to = pitch_statsbomb)

passes_rescaled <- data.frame(x  = to_statsbomb$x(filtered_passes$LocationX),
                              y  = to_statsbomb$y(filtered_passes$LocationY),
                              x2 = to_statsbomb$x(filtered_passes$EndLocationX),
                              y2 = to_statsbomb$y(filtered_passes$EndLocationY))

ggplot(passes_rescaled) +
  annotate_pitch(dimensions = pitch_statsbomb, , colour = "white", fill   = "springgreen4") +
  geom_segment(aes(x = x, y = y, xend = x2, yend = y2),
               colour = "coral",
               arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
   geom_point(aes(x = x, y = y),
             colour = "yellow",
             size = 4) +
  theme_pitch() +
  direction_label(x_label = 60) +
  theme(panel.background = element_rect(fill = "springgreen4"))+
  ggtitle("PassMap",
          "Messi in the First Half (Normal time)")
```

## Second Half (Normal time)

```{r}
filtered_passes <- process_period(2)
to_statsbomb <- rescale_coordinates(from = pitch_opta, to = pitch_statsbomb)

passes_rescaled <- data.frame(x  = to_statsbomb$x(filtered_passes$LocationX),
                              y  = to_statsbomb$y(filtered_passes$LocationY),
                              x2 = to_statsbomb$x(filtered_passes$EndLocationX),
                              y2 = to_statsbomb$y(filtered_passes$EndLocationY))

ggplot(passes_rescaled) +
  annotate_pitch(dimensions = pitch_statsbomb, , colour = "white", fill   = "springgreen4") +
  geom_segment(aes(x = x, y = y, xend = x2, yend = y2),
               colour = "coral",
               arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
   geom_point(aes(x = x, y = y),
             colour = "yellow",
             size = 4) +
  theme_pitch() +
  direction_label(x_label = 60) +
  theme(panel.background = element_rect(fill = "springgreen4"))+
  ggtitle("PassMap",
          "Messi in the Second Half (Normal time)")
```

## First Half (Extra time)

```{r}
filtered_passes <- process_period(3)

to_statsbomb <- rescale_coordinates(from = pitch_opta, to = pitch_statsbomb)

passes_rescaled <- data.frame(x  = to_statsbomb$x(filtered_passes$LocationX),
                              y  = to_statsbomb$y(filtered_passes$LocationY),
                              x2 = to_statsbomb$x(filtered_passes$EndLocationX),
                              y2 = to_statsbomb$y(filtered_passes$EndLocationY))

ggplot(passes_rescaled) +
  annotate_pitch(dimensions = pitch_statsbomb, , colour = "white", fill   = "springgreen4") +
  geom_segment(aes(x = x, y = y, xend = x2, yend = y2),
               colour = "coral",
               arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
   geom_point(aes(x = x, y = y),
             colour = "yellow",
             size = 4) +
  theme_pitch() +
  direction_label(x_label = 60) +
  theme(panel.background = element_rect(fill = "springgreen4"))+
  ggtitle("PassMap",
          "Messi in the First Half (Extra time)")
```

## Second Half (Extra time)

```{r}
filtered_passes <- process_period(4)

to_statsbomb <- rescale_coordinates(from = pitch_opta, to = pitch_statsbomb)

passes_rescaled <- data.frame(x  = to_statsbomb$x(filtered_passes$LocationX),
                              y  = to_statsbomb$y(filtered_passes$LocationY),
                              x2 = to_statsbomb$x(filtered_passes$EndLocationX),
                              y2 = to_statsbomb$y(filtered_passes$EndLocationY))

ggplot(passes_rescaled) +
  annotate_pitch(dimensions = pitch_statsbomb, , colour = "white", fill   = "springgreen4") +
  geom_segment(aes(x = x, y = y, xend = x2, yend = y2),
               colour = "coral",
               arrow = arrow(length = unit(0.25, "cm"),
                             type = "closed")) +
   geom_point(aes(x = x, y = y),
             colour = "yellow",
             size = 4) +
  theme_pitch() +
  direction_label(x_label = 60) +
  theme(panel.background = element_rect(fill = "springgreen4"))+
  ggtitle("PassMap",
          "Messi in the Second Half (Extra time)")
```

## Pass Position

### Average

```{r}
cleanevents %>%
  filter(player.id==player_id & type.name == "Pass") %>%
  soccerTransform(method='statsbomb') %>%
  soccerPositionMap(id = "player.name", x = "location.x", y = "location.y",
                    fill1 = "blue",
                    arrow = "r",
                    theme = "grass",
                    title = "Average Pass Postition",
                    subtitle='Messi vs France')

```

### Favourite

```{r}
 cleanevents %>%
  filter(player.id == player_id) %>%
  soccerPassmap(fill = "lightblue", arrow = "r", theme='grass',
                title = "Highest Pass Position")
```

## Comparison

```{r}
cleanevents %>%
  filter(team.name=='Argentina' & type.name == "Pass") %>%
  soccerTransform(method='statsbomb') %>%
  soccerPositionMap(id = "player.name", x = "location.x", y = "location.y",
                    fill1 = "blue",
                    arrow = "r",
                    theme = "grass",
                    title = "Average Pass Postition",
                    subtitle='Argentina')

```

# Shots

## Comparison among teams

```{r}
summary <- cleanevents %>%
  group_by("Team"=cleanevents$team.name) %>%
  summarise(shots = sum(type.name=="Shot", na.rm = TRUE))
summary
```

## Comparison among Argentinian players

```{r}
filtered_events <- cleanevents %>%
  filter(team.name == 'Argentina' & type.name == 'Shot') %>%
  select(player.name)

shots <- filtered_events %>%
  group_by(player.name) %>%
  summarise(shots = n())

print(shots)
```

## Messi's ShotMap

-   23' (P)
-   108'
-   Shootouts

```{r}
cleanevents %>%
  filter(player.id==player_id) %>%
  soccerTransform(method='statsbomb') %>%
  soccerShotmap(theme = "grass", title = "ShotMap",
                subtitle = "Shots taken by Messi")
```

## Messi's Average ShotMap

```{r}
cleanevents %>%
  filter(player.id==player_id & type.name == "Shot") %>%
  soccerTransform(method='statsbomb') %>%
  soccerPositionMap(id = "player.name", x = "location.x", y = "location.y",
                    fill1 = "blue",
                    arrow = "r",
                    theme = "grass",
                    title = "Average Shot Postition",
                    subtitle = 'Messi vs France')
```

# Movement

## HeatMap

```{r}
cleanevents %>%
  filter(player.id==player_id) %>%
  soccerTransform(method='statsbomb') %>%
  soccerHeatmap(x = "location.x", y = "location.y", xBins = 21, yBins = 14,
                title = "HeatMap",
                subtitle = 'Messi vs France')
```

# Model

## Pre-processing

```{r}
data <- cleanevents %>%
  select(duration, player.name)
```

```{r}
filtered_events_model <- cleanevents %>%
  filter(team.name=='Argentina') %>%
  select(duration, player.name)

data_clean <- filtered_events_model[complete.cases(filtered_events_model$duration, filtered_events_model$player.name), ]

print(data_clean)
```

```{r}
duration_data <- data.frame(total_duration = double(), player_name = character(), stringsAsFactors = FALSE)
```

```{r}
process_period <- function(player_name) {
  filtered_events <- data_clean %>%
  filter(player.name==player_name) %>%
  select(duration)

  total_duration_player <- sum(filtered_events$duration, na.rm = TRUE)

  new_row <- data.frame(total_duration = total_duration_player, player_name = player_name, stringsAsFactors = FALSE)
  return (new_row)
}
```

```{r}
unique_players <- unique(data_clean$player.name)


for (player_name in unique_players) {
  new_row <- process_period(player_name)
  duration_data <- rbind(duration_data, new_row)

}

print(duration_data)
```

```{r}
#Total shots by each player
shots <- cleanevents %>%
  filter(team.name == 'Argentina' & type.name == 'Shot') %>%
  group_by(player.name) %>%
  summarise(shots = n())

# Create a data frame with all player names
all_players <- data.frame(player.name = unique(cleanevents$player.name[cleanevents$team.name == 'Argentina']))

# Left join with shots summary, replace NA with 0
merged_data <- left_join(all_players, shots, by = "player.name") %>%
  mutate(shots = ifelse(is.na(shots), 0, shots))

# Print the result
print(merged_data)

```

```{r}
# Assuming merged_data and duration_data are your data frames
# merged_data contains columns player.name and shots
# duration_data contains columns player_name and total_duration

# Left join merged_data with duration_data
final_merged_data <- left_join(merged_data, duration_data, by = c("player.name" = "player_name"))

# If needed, replace NA values in shots and total_duration with 0
final_merged_data$shots[is.na(final_merged_data$shots)] <- 0
final_merged_data$total_duration[is.na(final_merged_data$total_duration)] <- 0

# Print the final merged data
print(final_merged_data)


```

```{r}
# Filter and count passes
passes <- cleanevents %>%
  filter(team.name == 'Argentina' & type.name == 'Pass') %>%
  group_by(player.name) %>%
  summarise(passes = n())

# Merge with the existing final_merged_data
final_merged_data <- left_join(final_merged_data, passes, by = c("player.name" = "player.name"))

# If needed, replace NA values in passes with 0
final_merged_data$passes[is.na(final_merged_data$passes)] <- 0

# Print the final merged data
print(final_merged_data)

```

## k-means Clustering

```{r}
selected_columns <- final_merged_data[, c("shots", "total_duration", "passes")]

# Standardize the data
scaled_data <- scale(selected_columns)

k <- 3  
kmeans_result <- kmeans(scaled_data, centers = k, nstart = 20)

# Add the cluster assignments back to the original data frame
final_merged_data$cluster <- kmeans_result$cluster

# Print the cluster centroids
print(kmeans_result$centers)

# View the distribution of players in each cluster
table(final_merged_data$cluster)

```

```{r}
# Scatter plot of total_duration against shots, colored by cluster, with player names
ggplot(final_merged_data, aes(x = total_duration, y = shots, color = factor(cluster), label = player.name)) +
  geom_point(aes(size = passes)) +
  geom_text(nudge_x = 0.2, nudge_y = 0.2, size = 3) +  # Adjust nudge values as needed
  labs(title = "K-Means Clustering of Players", x = "Total Duration", y = "Shots", color = "Cluster") +
  theme_minimal()

# Draw circles around each cluster
cluster_centers <- as.data.frame(kmeans_result$centers[, c("total_duration", "shots")])
```

```{r}

ggplot(final_merged_data, aes(x = passes, y = shots, color = factor(cluster), label = player.name)) +
  geom_point(aes(size = total_duration)) +
  geom_text(nudge_x = 0.2, nudge_y = 0.2, size = 3) +  # Adjust nudge values as needed
  labs(title = "K-Means Clustering of Players", x = "Passes", y = "Shots", color = "Cluster") +
  theme_minimal()


cluster_centers <- as.data.frame(kmeans_result$centers[, c("passes", "shots")])
```

```{r}
library(plotly)

# Create a 3D scatter plot
p <- plot_ly(final_merged_data, x = ~total_duration, y = ~shots, z = ~passes, color = ~factor(cluster)) %>%
  add_markers(size = 2)  %>%

# Add labels for player names
  add_text(text = ~player.name, 
           showlegend = FALSE,
           hoverinfo = "text",
           size = 1)
# Show the plot
p



```
